---
title: "p8105_hw2_ml4420"
author: Mengjia Lyu
date: 2019-9-26
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

## Problem 1



```{r}


trash_data = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A:N"), 
                        col_names = TRUE, skip = 1) %>%
                        janitor::clean_names() %>%
                        na.omit()
# omit rows that do not include dumpster-specific data


# round the number of sports balls to the nearest integer
trash_data$sports_balls = as.integer(round(trash_data$sports_balls))

```

```{r}

# read and clean precipitation data for 2017 and 2018
precip_data_2017 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", col_names = TRUE, skip = 1) %>%
                              janitor::clean_names() %>%
                              na.omit() %>%
                              mutate(year = 2017)
 
precip_data_2018 = read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", col_names = TRUE, skip = 1) %>%
                              janitor::clean_names() %>%
                              na.omit() %>%
                              mutate(year = 2018)

# combine the two datasets
precip_data = full_join(precip_data_2017, precip_data_2018, by = NULL) %>%
                        mutate(month = month.name[month])
```

## Comments
The number of observations in the Mr. Trash Wheel dataset is `r nrow(trash_data)`. The number of observations in the precipitation dataset is `r nrow(precip_data)`. Key variables in the Mr. Trash Wheel dataset include `r names(trash_data)`. Key variables in the precipitation dataset include `r names(precip_data)`. Total precipitation in 2018 is `r sum(precip_data_2018$total)`. The median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trash_data, trash_data$year == 2017), sports_balls))`.


## Problem 2
```{r problem 2 part 1}
pols_month_data = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>%
                  janitor::clean_names() %>%
                  separate(mon, into = c("year", "month", "day"), sep = "-", convert = TRUE) %>%    #break up mon into integer variables
                  mutate(month = month.name[month]) %>%                                             #replace month number with month name
                  mutate(president = ifelse(prez_dem == 0, "gop", "dem")) %>%                       #create new variable president
                  select(-prez_dem) %>%                                                             #remove prez_dem, prez_gop, day
                  select(-prez_gop) %>%
                  select(-day)

```

```{r problem 2 part 2}
snp_data = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>%
           janitor::clean_names() %>%
           separate(date, into = c("month", "day", "year"), sep = "/", convert = TRUE) %>%          #arrange according to year and month
           mutate(month = month.name[month]) %>% 
           select(year, month, everything()) %>%
           select(-day)

```

```{r problem 2 part 3}
unemployment_data = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
                    janitor::clean_names() 
                    
  
colnames(unemployment_data) = c("year", "January", "February", "March", "April", "May", "June",
                                   "July", "August", "September", "October", "November", "December")

  
unemployment_tidy_data = 
  pivot_longer(
    unemployment_data,
    January:December,
    names_to = "month",
    values_to = "percentage_of_unemployment")

#merge snp into pols
snp_pols_data = 
  left_join(pols_month_data, snp_data, by = c("year", "month") )

#merge unemployment into the result
snp_pols_unemploy_data = 
  left_join(snp_pols_data, unemployment_tidy_data, by = c("year", "month"))

```
## Comments
**pols_month_data** contains 822 observations of 7 variables related to the number of senators, governors and representatives who are democratic or republican at any given time.
**snp_data** contains 787 observations of the date and the closing values of the Standard & Poor's stock index on the associated date.
**unemployment_tidy_data** and **unemployment_data** contains 68 observations of the percentage of unemployment in each month of the associated year.
**snp_pols_data** contains the observations from both **pols_month_data** and **snp_data**.
**snp_pols_unemploy_data** contains the observations from all of **pol_month_data**, **snp_data** and **unemployment_tidy_data**.

The resulting dataset has `r nrow(snp_pols_unemploy_data)` rows and `r ncol(snp_pols_unemploy_data)` columns.
Range of years is from `r min(snp_pols_unemploy_data$year)` to `r max(snp_pols_unemploy_data$year)`.
Names of key variables include `r names(snp_pols_unemploy_data)`.

## Problem 3
```{r}

library(tools)
baby_names_data = read_csv("./data/Popular_Baby_Names.csv") %>%
                  janitor::clean_names() %>%
                  mutate(gender = str_replace(gender, "FEMALE", "Female")) %>%
                  mutate(gender = str_replace(gender, "MALE", "Male"))

# change names to proper format
baby_names_data$childs_first_name = toTitleCase(tolower(baby_names_data$childs_first_name))

# change ethnicity to proper format
baby_names_data$ethnicity = str_replace_all(baby_names_data$ethnicity, "PACI$", "PACIFIC ISLANDER")
baby_names_data$ethnicity = str_replace_all(baby_names_data$ethnicity, "HISP$", "HISPANIC")
baby_names_data$ethnicity = toTitleCase(tolower(baby_names_data$ethnicity))

# remove duplicate rows
distinct(baby_names_data)                                                   
```

## Plotting

```{r}
library(arsenal)

olivia_data = baby_names_data[which(baby_names_data$childs_first_name == "Olivia"), ] %>%
              select(-gender) %>%
              select(-childs_first_name) %>%
              select(-count) %>%
              distinct()

olivia_tidy_data = pivot_wider(
  olivia_data,
  names_from = year_of_birth,
  values_from = rank            # rank in popularity of the name "Olivia" over time    
)

# create a reader-friendly table

knitr::kable(olivia_tidy_data)
          
# most popular name among male children over time
most_ppl_male_name_data = baby_names_data[which(baby_names_data$gender == "Male" &
                                                baby_names_data$ethnicity == "White Non Hispanic" &
                                                baby_names_data$year_of_birth == 2016), ] %>%
                                          select(-year_of_birth) %>%
                                          select(-gender) %>%
                                          select(-ethnicity) %>%
                                          select(-count)


```